---
title: "Lesson 10 Poisson regression"
author: "Chun Hsien Wu"
date: "2021年6月28日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Lesson 10.2
#### Data
For an example of Poisson regression, we’ll use the `badhealth` data set from the `COUNT` package in `R`.

```{r}
library("COUNT")
```

```{r}
data("badhealth")
head(badhealth)
```
It's also a good idea to check. To make sure we don't have any missing values. We are going to check using `is.na`, which will flag missing values, and we want to see if there any of those.
```{r}
any(is.na(badhealth))
```
Let's visidual the data
```{r}
hist(badhealth$numvisit, breaks =20)
```
As we can see, number of visits is positively valued or at least non-negative value, right skewed quantity.
Find the 0 number of value.
```{r}
min(badhealth$numvisit)
```
Let's count how many zeros are there
```{r}
sum(badhealth$numvisit==0)
```
```{r}
plot(jitter(log(badhealth$numvisit)) ~ jitter(age), data=badhealth, subset=badhealth$numvisit=0, xlab="age", ylab="log(visits)")
```

