---
title: "bayesian linear regression"
author: "Chun Hsien Wu"
date: "2021年6月3日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 貝氏迴歸分析使用R 及 car package 

使用 `car` 套件中的 `Leinhardt` 資料庫 

<http://rmarkdown.rstudio.com>.

 **Knit** 

```{r cars}
library("car")
data("Leinhardt")
?Leinhardt
head(Leinhardt)
```
使用  `str`function 了解 `Leinhardt` 資料庫有哪些資料類別

```{r}
str(Leinhardt)
```

```{r}
pairs(Leinhardt)
```

infant mortality and income doesn't like a linear relationshiop

```{r}
plot(infant ~ income, data=Leinhardt)
```
In fact, both of these variables are extremely right skewed(Matthew Heiner). Let's look at a histogram of each one individually.

```{r}
hist(Leinhardt$infant)
```

```{r}
hist(Leinhardt$income)
```

A linear regression is not appropriate for these variables.
We could try looking at this on the log scale. Let's recreate these plots now with variables that are transformed.

```{r}
Leinhardt$loginfant = log(Leinhardt$infant)
Leinhardt$logincome = log(Leinhardt$income)
plot(loginfant ~ logincome, data=Leinhardt)
```

## Modeling

## Bayesian Linear Regression using jags

以下JAGS 來分析資料

# Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
