---
title: "Lesson 7 Part A"
author: "Chun Hsien Wu"
date: "2021年6月8日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Anscombe data set bayesian analysis using JAGS

#### Question 3 
Consider the Anscombe data set in R
Suppose we are interested in relating per-capita education expenditures to the other three variables. Which variable appears to have the strongest linear relationship with per-capita education expenditures?

```{r}
library("car")  # load the 'car' package
data("Anscombe")  # load the data set
?Anscombe  # read a description of the data
head(Anscombe)  # look at the first few lines of the data
pairs(Anscombe)  # scatter plots for each pair of variables
```

```{r}
lmod = lm(education ~ income + young + urban, data=Anscombe)
summary(lmod)
```
The answer is **Per-capita** generally co -coour with increases in education expenditures.

#### Question 4
Fit a reference (noninformative) Bayesian linear model to the Anscombe data with education expenditures as the response variable and include all three other variables as predictors. Use the lm function in R.

What is the posterior mean estimate of the intercept in this model? Round your answer to one decimal place.

```{r}
lmod = lm(education ~ income + young + urban, data=Anscombe)
summary(lmod)
```

The answer is **-2.868**

## Including Plots
You can also embed plots, for example:
```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
